---
title: "STAT251MittenHmwk10"
author: "Ethan Mitten"
date: "6/16/2020"
output:
  word_document: default
  html_document: default
---

```{r}
library(readxl)
survey <- read_excel("Class survey (Five semester combined).xlsx")
```

**Using set.seed(12345), write a function ‘ BootstrapConfidenceInterval’ with parameters (data, R, confidence level) that returns bootstrap confidence interval. Make the default of confidence level be 95%.  The following steps will lead you to write such function.**

```{r}
sleep <- survey$`Hours of Sleep`
set.seed(12345)
BootstrapConfidenceInterval <- function(data, R, confidencelevel) {
  a <- matrix(replicate(R, sample(data, 50, replace=T)), byrow=F, ncol=R)
  amean <-colMeans(a)
  b <- quantile(amean, probs= confidencelevel)
}
partfive <- c(.025, .975)
```

**(a)	What is 95% confidence interval of the mean sleeping hour of Truman students?**

```{r}
print(BootstrapConfidenceInterval(sleep, 1000, partfive))
```

The 95% confidence interval of the mean sleeping hour of Truman students is 6.74975 to 7.28025.


**(b)	What is 99% confidence interval of the mean sleeping hour of Truman students?**

```{r}
partb <- c(.01, .99)
print(BootstrapConfidenceInterval(sleep, 1000, partb))
```

The 99% confidence interval of the mean sleeping hour of Truman students is 6.67792 to 7.36. 


**Using set.seed(12345), write a function ‘ Bootstraptest’ with parameters (data, R, mu0, alternative) that returns bootstrap p-value. The function should include all possible alternatives, right tailed, left tailed, and tow-sided test.**


**Test if Truman students sleeps more than 7 hours a day on the average.**
```{r}
set.seed(12345)
Bootstraptest <- function(data, R, mu0, alternative= c("right", "left", "twosided")) {
  b <- matrix(replicate(R, sample(data, 50, replace=T)), byrow=F, ncol=R)
  bmean = colMeans(b)
  if (alternative== "right")
    x <- sum(bmean>mu0)
  print(x/R)
  if (alternative== "left")
    d <- sum(bmean<mu0)
  if (alternative=="twosided")
    z <- sum(bmean!=mu0)
  
}
Bootstraptest(sleep, 1000, 7, "right")
```

The p-value of the bootstrap test is .531 meaning that there is not a significant difference for Truman students sleeping greater than 7 hours a day on average.


**) Test if Truman female students sleeps less than 7 hours a day on the average.**
```{r}
set.seed(12345)
Bootstraptest <- function(data, R, mu0, alternative= c("right", "left", "twosided")) {
  b <- matrix(replicate(R, sample(data, 50, replace=T)), byrow=F, ncol=R)
  bmean = colMeans(b)
  if (alternative== "left")
    d <- sum(bmean<mu0)
  print(d/R)
  if (alternative=="twosided")
    z <- sum(bmean!=mu0)
  
}
Bootstraptest(sleep[survey$Sex=='f'], 1000, 7, "left")
```

The p-value of the bootstrap test is .574 meaning that there is not a significant difference for Truman female students to sleeping less than 7 hours a day on average.


