---
title: "STAT251MittenHmwk6"
author: "Ethan Mitten"
date: "6/9/2020"
output:
  word_document: default
  html_document: default
---

```{r}
library(readxl)
stress <- read_excel("~/Desktop/Junior Summer/STAT 251/stress.xlsx")
```


**(a)	Plot a scatterplots between Y and Xs. Is there any potential relationship observed?**
```{r}
plot(stress~`firm size`+ `years in position` + salary + age, data = stress)
```

Based on the scatterplots it can be seen that all of the variables seem to have a potentially positive relationship with stress except for age. 

**(b)	Check all the inference assumptions. Are there any violation?**
```{r}
lslmodel <- lm(stress~`firm size`+ `years in position` + salary + age, data = stress)
layout(matrix(c(1,2,3,4),2,2))
plot(lslmodel)
```

From the plots there does not seem to be any violations.


**(c)	Find the least square line.**
```{r}
lslmodel <- lm(stress~`firm size`+ `years in position` + salary + age, data = stress)
summary(lslmodel)
```

**(d)	Estimate all coefficients using 95% confidence intervals.**
```{r}
lslmodel$coef
```

**(e)	Are stress level and the amount of salary significantly related when all other variables were already considered?**
```{r}
summary(lslmodel)
```

Looking at the summary of the regression model it can be seen that with the salary line the estimate is at 2.1 and the p-value is 1.77e-06. Based on a 95% significance level we would say that these variables are significantly related.


**(f)	Using stepwise AIC find the best model.**
```{r}
step(lslmodel)
```

Using the stepwise AIC it looks like the best model involves all of the variables except years in position.


**(g)	What is the coefficient of determination for the model chosen from (f)? Interpret it.**
```{r}
determine <- lm(stress~`firm size` + salary + age, data = stress)
summary(determine)
```

The coefficient of determination is .6767. This is explaining the proportion of the total variation that is being explained by the regression line, and because we have a coefficient of determination of .6767 we would say this is a fairly good fit as it is closer to 1 than to 0.


























