---
title: "STAT 251 Assignment 5"
author: "Ethan Mitten"
date: "6/2/2020"
output:
  word_document: default
  html_document: default
---

```{r}
library(readxl)
mtcars <- read_excel("mtcars.xlsx")
attach(mtcars)
```

```{r}
Q1 <- t.test(mtcars$mpg, conf.level = .95)
```

```{r}
Q1$conf.int
```
There is a 95% chance that the true mean of the column mpg falls in the range of 17.91768 and 22.26357.

```{r}
t.test(mpg[cyl=='8'], conf.level = .95)
```
There is a 95% chance that the true mean of mpg for 8-cylinder cars in the data is between 13.62187 and 16.57813.

```{r}
t.test(mpg[vs=='1'], mu=20, alternative = "greater")
```
Based off the p-value being .003691 we would reject the null hypothesis because .003691 < .05. This would mean that based on our test the true mean of mpg when vs is equal to 1 is greater than 20.

```{r}
t.test(mpg[hp>'200'], mpg[hp<=200])
```
Based on the test and the p-value obtained from the test the mpg is not signficantly different for cars under 200 horsepower in comparison to cars above 200 horsepower. This is confirmed because the p-value of .6566 > .05.

```{r}
library(readxl)
survey <- read_excel("Class survey (Seven semester combined).xlsx")  
survey$`Shoe Size` <- as.numeric(as.character(survey$`Shoe Size`))
```

```{r}
shoesize <- survey$`Shoe Size`
handed <- survey$Handed
t.test(shoesize[handed=='L'], shoesize[handed=='r'])
```
Based on the test and the p-value obtained from the test the shoe size is not signficantly different based on whether you are right or left handed. This is confirmed because the p-value of .3038 > .05.

```{r}
sleep <- survey$`Hours of Sleep`
dogcat <- c(survey$DogsCats)
hist(sleep[dogcat=='cat'], ylim= c(0,100), col = "purple", main = "Hours of Sleep for Students Who Own a Cat", xlab = "Hours of Sleep", ylab = "Number of Students")
hist(sleep[dogcat=='dog'], ylim = c(0,100), col = "darkgreen", main = "Hours of Sleep for Students Who Own a Dog", xlab = "Hours of Sleep", ylab = "Number of Students")
t.test(sleep[dogcat=='cat'], sleep[dogcat=='dog'])
```
In this test I want to see if the hours of sleep a student gets is based on whether or not the student has a dog or a cat. Based on the test and the p-value obtained from the test the hours of sleep seems to be signficantly different based on whether you own a cat or a dog. This can be confirmed because the p-value of .0104 <.05.









































